---
title: "Transform Data"
output: html_notebook
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="")
library(tidyverse)
# install.packages("babynames")
library(babynames)
```


# dplyr
```{r}
ggplot(data = babynames) + 
  geom_line(aes(x = year, y = prop))
```

## Your Turn 1

Alter the code to select just the `n` column:

```{r}
select(babynames, n)
```
can do a range with a :
can do everything except with -
can do everything that starts with c with starts_with("c")
  or ends with
contains
one_of
num_range
    
## Quiz

Which of these is NOT a way to select the `name` and `n` columns together?
 ends with would not include name because name does not end with n

```{r}
select(babynames, -c(year, sex, prop))
select(babynames, name:n)
select(babynames, starts_with("n"))
select(babynames, ends_with("n"))
```

## Your Turn 2

Use `filter`, `babynames`, and the logical operators to find:

* All of the names where prop is greater than or equal to 0.08  
* All of the children named "Sea"  

```{r}
filter(babynames, name=="Garrett")

```

## Your Turn 3

Use Boolean operators to return only the rows that contain:

* _Boys_ named Sue  
* Names that were used by exactly 5 or 6 children in 1880  
* Names that are one of Acura, Lexus, or Yugo

```{r}
filter(babynames, sex=="M"& name=="Sue")
filter(babynames, xor(5,6))
```

## Help Me

What is the smallest value of n?
What is the largest?

```{r}

```


***

# Take aways

* Extract variables with `select()`  
* Extract cases with `filter()`  
* Arrange cases, with `arrange()`  


